\chapter*{}
\addcontentsline{toc}{chapter}{Acknowledgements}

\begin{center}
\vspace*{0.8cm}
{\Huge\bfseries Acknowledgements}
\end{center}

\vspace{0.8cm}

I would like to express my sincere gratitude to my supervisor,
\textbf{Dr.\ T.\ K.\ Shajahan}, for his guidance, encouragement,
and continuous support throughout this project. His valuable
suggestions and insightful discussions have been instrumental in
shaping this work.

\smallskip

I am deeply grateful to \textbf{Vivek B.\ Meshram}, Ph.D. Scholar,
Department of Physics, NITK Surathkal, for his constant assistance,
technical guidance, and helpful discussions throughout this work.
His support in understanding the underlying concepts and resolving
technical difficulties was invaluable to the successful completion
of this project.

\smallskip

I would like to thank my project group members,
\textbf{Deepanshu Kurmi} and \textbf{Sribarna Haldar}, for their
valuable discussions, cooperation, and support during the course of
this work.

\smallskip

I am grateful to the faculty members and staff of the
\textbf{Department of Physics, NITK Surathkal}, for providing an
excellent academic environment and the facilities necessary for
carrying out this research.

\smallskip

Finally, I express my deepest gratitude to my parents and family
for their unwavering support, patience, and encouragement
throughout my studies. Their constant motivation has been a source
of strength and inspiration.

\vspace{2cm}

\begin{flushright}
\textbf{Pranav D. Shinde}
\end{flushright}
